### 2020-02-03 Supplement_forSupershort.Rmd
cleanup of previous submission materials and versions, organized to move forward with PlosBio/PNAS version

---
bibliography: manuscript.bib
csl: systematic-biology.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
language:
  label:
    fig: "**Figure S**"
    tab: "**Table S**"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \doublespacing
---

@@

@@

@@

![\label{appendix2} Character definitions. ](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/Appendix2.png){height=500px}

@@

![\label{MLIqtree} Maximum likelihood IQTree inference, unconstrained. Node labels are bootstrap support values.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/MLunconstrained.png){height=300px}

@@

![\label{Transcriptome_constraint} Topology used to constrain analyses (minimal topological statements based on the incongruences between the unconstrained tree and Munro et al. (2018). ](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/constrain.png){height=400px}

![\label{ML_constrained} Constrained IQTree ML inference. Node labels are bootstrap support values.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/MLconstrained.png){height=300px}

![\label{Bayes_unconstrained} Unconstrained Bayesian topology inference in RevBayes (node labels are Bayesian posteriors).](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/BAYESunconstrained.png){height=300px}

![\label{Bayes_constrained} Clade constrained Bayesian inference in RevBayes (node labels are Bayesian posteriors).](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/BayesConstrained.png){height=300px}

![\label{unconstrained_timetree} Unconstrained ultrametric Bayesian time tree branch length and topology inference in RevBayes (node labels are Bayesian posteriors). Arbitrary rooting.).](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/Unconstrained timetree.png){height=400px}

![\label{constrained_timetree} Ultrametric Bayesian time tree branch length inference in RevBayes (node labels are bayesian posteriors). Topology clamped to the Bayesian constrained topology inference in \@ref{Bayes_constrained}. Tree rooted using outgroup constraint.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/constrainedTimetree.png){height=500px}

@@

![\label{ModelSupport} Model support (delta AICc), phylogenetic signal (Blomberg’s K), and phylogenetic signal permutation test p-value for each continuous character. Ntaxa = Number of taxa used in the analyses after removing those where the character sate is inapplicable or the data is missing.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/modelSupport.png){height=600px}

![\label{ModelAdequacy} Model adequacy scores for the best model supported for each morphological character. Cvar = coefficient of variation of the absolute value of the contrasts. Svar = Slope of a linear model fitted to the absolute value of the contrasts against their expected variances. Sasr = slope of the contrasts against the ancestral state inferred at each corresponding node. Shgt = slope of the contrasts against node depth. Dcfd = Kolmolgorov-Smirnov D-statistic comparing contrasts to a normal distribution with SD equal to the root of the mean of squared contrasts. ](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/modelAdequacy.png)

![\label{ModelSupportOUwie} Model support (delta AICc) for each morphological character analyzed on the feeding guild reconstruction regime tree. OU1 = Single-optimum Ornstein-Uhlenbeck. OUm = Multi-optima Ornstein-Uhlenbeck. Model adequacy scores calculated for the best supported model only. Msig = mean of squared contrasts. Cvar = coefficient of variation of the absolute value of the contrasts. Svar = Slope of a linear model fitted to the absolute value of the contrasts against their expected variances. Sasr = slope of the contrasts against the ancestral state inferred at each corresponding node. Shgt = slope of the contrasts against node depth. Dcfd = Kolmolgorov-Smirnov D-statistic comparing contrasts to a normal distribution with SD equal to the root of the mean of squared contrasts.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/OUwie_modelsupport.png)

@@

![\label{DAPCguilds} DAPC for Feeding guilds. Six PCs retained after a-score optimization (100 iterations). Four LDA functions used. Discriminant power on training set: 100%. Prediction posterior distribution heat map in main text Figure 6. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/DAPC_guilds.png){height=500px}

![\label{DAPCcopepods} DAPC for copepod presence in the diet. Eight PCs retained after a-score optimization (100 iterations). One LDA functions used. Discriminant power on training set: 95.4%. Grayscale heat map shows the posterior probability distribution of the predictions. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/DAPC_copepod.png)

![\label{DAPCfish} DAPC for fish presence in the diet. Three PCs retained after a-score optimization (100 iterations). One LDA function used. Discriminant power on training set: 68.1%. Grayscale heat map shows the posterior probability distribution of the predictions. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/DAPC_fish.png)

![\label{DAPCcrustacean} DAPC for large crustacean presence in the diet. Four PCs retained after a-score optimization (100 iterations). One LDA function used. Discriminant power on training set: 81.8%. Grayscale heat map shows the posterior probability distribution of the predictions. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/DAPC_crustacean.png)

@@

![\label{GLMs} Logistic regressions between continuous morphological characters and prey type presences. Ntaxa = number of taxa used in the analyses after removing taxa with missing diet data and inapplicable character states. phyloGLM = Phylogenetic generalized logistic regression model. GLM = Generalized logistic regression model. P = p-value. b = slope. Only cases with significant GLM fits were retained. Cells colored blue indicate phyloGLM p-value < 0.05. Cells colored green indicate GLM p-value < 0.05](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/PhyloGLM.png){height=300px}

![\label{SIMMAPguilds} SIMMAP Feeding guilds.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/guild_SIM.png)

@@

![\label{VCV_NAZero} Rate covariance matrix for the whole tree using all taxa (45 species), transforming inapplicable states to zeroes. Covariances scaled to correlations. All characters estimated simultaneously under Brownian Motion.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCV_Nazero.png)

![\label{VCV_NAremove} Rate covariance matrix for the whole tree using only taxa without inapplicable states (24 species). Covariances scaled to correlations. All characters estimated simultaneously under Brownian Motion.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCV_NAremove.png)

![\label{VCV_dietaxa} Rate covariance matrix for the whole tree using only taxa with diet data (22 species), transforming inapplicable states to zeroes. Covariances scaled to correlations. All characters estimated simultaneously under Brownian Motion.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCV_dietaxa.png)

![\label{VCV_bestmodels} Best models (lowest AIC) supported in a pairwise character rate covariance analysis comparing correlated Brownian Motion models across the five selective regimes. Selective regimes were mapped onto the tree using an ancestral state reconstruction of the feeding guilds. Blank cells represent computationally singular contrasts.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCVbestModels.png)

![\label{VCV_ntaxa} Number of taxa used for each pairwise contrast in the VCV analyses, given the number of taxa without inapplicable states.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCV_ntaxa.png)

![\label{VCV_regimes} Pairwise estimated rate covariance matrices across the five selective regimes, using only taxa with diet data. Covariances scaled to correlations. Selective regimes were mapped onto the tree (22 species with diet data) using a stochastic mapping of the feeding guilds. Tree is pruned to taxa with no inapplicable states for a given character pair. Not all regimes are represented in all contrasts. Question marks represent computationally singular contrasts.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCV_regimes.png){height=600px}

![\label{VCV_diffs} Scaled differences between the regime-specific covariance matrices in \@ref{VCV_regimes} and the whole tree covariance matrix.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCVdiffs.png){height=600px}

![\label{VCV_diffs_anc} Scaled differences between the regime-specific covariance matrices in \@ref{VCV_regimes} and the covariance matrices in their preceding regime, the large-crustacean specialist regime.](Post-submission modifications/Supplementary_materials/Online_appendices/PNGs/VCVdiffs_anc.png)

>>>>

---
bibliography: manuscript.bib
csl: pnas.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
language:
  label:
    fig: "**Figure S**"
    tab: "**Table S**"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \doublespacing
---

@@

Phylogenetic inference: We aligned the sequences using MAFFT [@katoh2002mafft] (alignments available in Dryad). We inferred a Maximum Likelihood (ML) phylogeny (SM6) from 16S and 18S ribosomal rRNA genes using IQTree [@nguyen2014iq] with 1000 bootstrap replicates (iqtree -s alignment.fa -nt AUTO -bb 1000). We used ModelFinder [@kalyaanamoorthy2017modelfinder] implemented in IQTree v1.5.5. to assess the relative model fit. ModelFinder selected GTR+R4 for having the lowest Bayesian Information Criterion score. Additionally, we inferred a Bayesian tree with each gene as an independent partition in RevBayes [@hohna2016revbayes] (SM9 and SM11), which was topologically congruent with the unconstrained ML tree. The *alpha* priors were selected to minimize prior load in site variation.

Diet data curation: We removed the gelatinous prey observations for *Praya dubia* eating a ctenophore and a hydromedusa, and for *Nanomia* sp. eating *Aegina* since we believe these are rare events that have a much larger probability of being detected by ROV methods than their usual prey, and it is not clear whether the medusae were attempting to prey upon the siphonophores. Personal observations on feeding (from SHDH, CAC, and Philip Pugh) were also included for *Resomia ornicephala*, *Lychnagalma utricularia*, *Bargmannia amoena*, *Erenna richardi*, *Erenna laciniata*, *Erenna sirena*, and *Apolemia rubriversa*.

The feeding guilds decalred in this study are: small-crustacean specialist (feeding mainly on copepods and ostracods), large crustacean specialist (feeding on large decapods, mysids, or krill), fish specialist (feeding mainly on actinopterygian larvae, juveniles, or adults), gelatinous specialist (feeding mainly on other siphonophores, medusae, ctenophores, salps, and/or doliolids), and generalist (feeding on a combination of the aforementioned taxa, without favoring any one prey group). These were selected to minimize the number of categories while keeping the most different types of prey separate.

Data wrangling for comparative analyses: For comparative analyses, we removed species present in the tree but not represented in the morphology data, and *vice versa*. Although we measured specimens labeled as *Nanomia bijuga* and *Nanomia cara*, we are not confident in some of the species-level identifications, and some specimens were missing diagnostic zooids. Thus, we decided to collapse these into a single taxonomic concept (*Nanomia* sp.). All *Nanomia* sp. observations were matched to the phylogenetic position of *Nanomia bijuga* in the tree. We carried out all phylogenetic comparative statistical analyses in the programming environment R [@team2017r], using the Bayesian ultrametric species tree (Fig. \@ref(figure4)), and incorporating intraspecific variation estimated from the specimen data as standard error whenever the analysis tool allowed it. R scripts and summarized species-collapsed data available in the Dryad repository. For each character (or character pair) analyzed, we removed species with missing data and reported the number of taxa included. We tested each character for normality using the Shapiro-Wilk test [@shapiro1965analysis], and log-transformed those that were non-normal.

Data wrangling for the variance-covariance analyses: When fitting all variance-covariance terms simultaneously (SM36-38), we selected the largest set of characters that would allow the analysis to run without computational singularities. This excluded many of the morphometric characters which are linearly dependent on other characters. Since the functions do not tolerate missing data, we ran the analyses in two ways: One including all taxa but transforming absent states to zeroes, and another removing the taxa with absent states. These analyses could only be carried out on the subset of taxa for which diet data is available, and only among character pairs that are not computationally singular for that taxonomic subset. Gelatinous specialist correlations could only be estimated for a small subset of characters present in *Apolemia* and should be interpreted with care.

Comparative tools used to test character associations: To test for correlated evolution among binary characters, we used Pagel’s test [@pagel1994detecting]. To characterize and evaluate the relationship between continuous characters, we used phylogenetic generalized least squares regressions (PGLS) [@grafen1989phylogenetic]. To compare the evolution of continuous characters with categorical aspects of the diet, we carried out a phylogenetic logistic regression (R nlme::gls using the 'corBrownian' function for the argument ‘correlation’).

@@

The discriminant analysis for feeding guild (7 principal components, 4 discriminants) produced 100% discrimination, and the highest loading contributions were found for the characters (ordered from highest to lowest): Involucrum length, heteroneme volume, heteroneme number, total heteroneme volume, tentacle width, heteroneme length, total nematocyst volume, and heteroneme width (SM21).

@@

![\label{appendix2} Character definitions. ](Supplementary_materials/Online_appendices/PNGs/Appendix2.png){height=500px}

@@

![\label{MLIqtree} Maximum likelihood IQTree inference, unconstrained. Node labels are bootstrap support values.](Supplementary_materials/Online_appendices/PNGs/MLunconstrained.png){height=300px}

@@

![\label{Transcriptome_constraint} Topology used to constrain analyses (minimal topological statements based on the incongruences between the unconstrained tree and Munro et al. (2018). ](Supplementary_materials/Online_appendices/PNGs/constrain.png){height=400px}

![\label{ML_constrained} Constrained IQTree ML inference. Node labels are bootstrap support values.](Supplementary_materials/Online_appendices/PNGs/MLconstrained.png){height=300px}

![\label{Bayes_unconstrained} Unconstrained Bayesian topology inference in RevBayes (node labels are Bayesian posteriors).](Supplementary_materials/Online_appendices/PNGs/BAYESunconstrained.png){height=300px}

![\label{Bayes_constrained} Clade constrained Bayesian inference in RevBayes (node labels are Bayesian posteriors).](Supplementary_materials/Online_appendices/PNGs/BayesConstrained.png){height=300px}

![\label{unconstrained_timetree} Unconstrained ultrametric Bayesian time tree branch length and topology inference in RevBayes (node labels are Bayesian posteriors). Arbitrary rooting.).](Supplementary_materials/Online_appendices/PNGs/Unconstrained timetree.png){height=400px}

![\label{constrained_timetree} Ultrametric Bayesian time tree branch length inference in RevBayes (node labels are bayesian posteriors). Topology clamped to the Bayesian constrained topology inference in \@ref{Bayes_constrained}. Tree rooted using outgroup constraint.](Supplementary_materials/Online_appendices/PNGs/constrainedTimetree.png){height=500px}

@@

![\label{ModelSupport} Model support (delta AICc), phylogenetic signal (Blomberg’s K), and phylogenetic signal permutation test p-value for each continuous character. Ntaxa = Number of taxa used in the analyses after removing those where the character sate is inapplicable or the data is missing.](Supplementary_materials/Online_appendices/PNGs/modelSupport.png){height=600px}

![\label{ModelAdequacy} Model adequacy scores for the best model supported for each morphological character. Cvar = coefficient of variation of the absolute value of the contrasts. Svar = Slope of a linear model fitted to the absolute value of the contrasts against their expected variances. Sasr = slope of the contrasts against the ancestral state inferred at each corresponding node. Shgt = slope of the contrasts against node depth. Dcfd = Kolmolgorov-Smirnov D-statistic comparing contrasts to a normal distribution with SD equal to the root of the mean of squared contrasts. ](Supplementary_materials/Online_appendices/PNGs/modelAdequacy.png)

![\label{ModelSupportOUwie} Model support (delta AICc) for each morphological character analyzed on the feeding guild reconstruction regime tree. OU1 = Single-optimum Ornstein-Uhlenbeck. OUm = Multi-optima Ornstein-Uhlenbeck. Model adequacy scores calculated for the best supported model only. Msig = mean of squared contrasts. Cvar = coefficient of variation of the absolute value of the contrasts. Svar = Slope of a linear model fitted to the absolute value of the contrasts against their expected variances. Sasr = slope of the contrasts against the ancestral state inferred at each corresponding node. Shgt = slope of the contrasts against node depth. Dcfd = Kolmolgorov-Smirnov D-statistic comparing contrasts to a normal distribution with SD equal to the root of the mean of squared contrasts.](Supplementary_materials/Online_appendices/PNGs/OUwie_modelsupport.png)

@@

![\label{DAPCguilds} DAPC for Feeding guilds. Six PCs retained after a-score optimization (100 iterations). Four LDA functions used. Discriminant power on training set: 100%. Prediction posterior distribution heat map in main text Figure 6. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Supplementary_materials/Online_appendices/PNGs/DAPC_guilds.png){height=500px}

![\label{DAPCcopepods} DAPC for copepod presence in the diet. Eight PCs retained after a-score optimization (100 iterations). One LDA functions used. Discriminant power on training set: 95.4%. Grayscale heat map shows the posterior probability distribution of the predictions. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Supplementary_materials/Online_appendices/PNGs/DAPC_copepod.png)

![\label{DAPCfish} DAPC for fish presence in the diet. Three PCs retained after a-score optimization (100 iterations). One LDA function used. Discriminant power on training set: 68.1%. Grayscale heat map shows the posterior probability distribution of the predictions. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Supplementary_materials/Online_appendices/PNGs/DAPC_fish.png)

![\label{DAPCcrustacean} DAPC for large crustacean presence in the diet. Four PCs retained after a-score optimization (100 iterations). One LDA function used. Discriminant power on training set: 81.8%. Grayscale heat map shows the posterior probability distribution of the predictions. Variable contribution (top quartile) calculated by the sum of the LDA variable loadings weighted by the eigenvalue of each LDA.](Supplementary_materials/Online_appendices/PNGs/DAPC_crustacean.png)

@@

![\label{GLMs} Logistic regressions between continuous morphological characters and prey type presences. Ntaxa = number of taxa used in the analyses after removing taxa with missing diet data and inapplicable character states. phyloGLM = Phylogenetic generalized logistic regression model. GLM = Generalized logistic regression model. P = p-value. b = slope. Only cases with significant GLM fits were retained. Cells colored blue indicate phyloGLM p-value < 0.05. Cells colored green indicate GLM p-value < 0.05](Supplementary_materials/Online_appendices/PNGs/PhyloGLM.png){height=300px}

![\label{SIMMAPguilds} SIMMAP Feeding guilds.](Supplementary_materials/Online_appendices/PNGs/guild_SIM.png)

@@

![\label{VCV_NAZero} Rate covariance matrix for the whole tree using all taxa (45 species), transforming inapplicable states to zeroes. Covariances scaled to correlations. All characters estimated simultaneously under Brownian Motion.](Supplementary_materials/Online_appendices/PNGs/VCV_Nazero.png)

![\label{VCV_NAremove} Rate covariance matrix for the whole tree using only taxa without inapplicable states (24 species). Covariances scaled to correlations. All characters estimated simultaneously under Brownian Motion.](Supplementary_materials/Online_appendices/PNGs/VCV_NAremove.png)

![\label{VCV_dietaxa} Rate covariance matrix for the whole tree using only taxa with diet data (22 species), transforming inapplicable states to zeroes. Covariances scaled to correlations. All characters estimated simultaneously under Brownian Motion.](Supplementary_materials/Online_appendices/PNGs/VCV_dietaxa.png)

![\label{VCV_bestmodels} Best models (lowest AIC) supported in a pairwise character rate covariance analysis comparing correlated Brownian Motion models across the five selective regimes. Selective regimes were mapped onto the tree using an ancestral state reconstruction of the feeding guilds. Blank cells represent computationally singular contrasts.](Supplementary_materials/Online_appendices/PNGs/VCVbestModels.png)

![\label{VCV_ntaxa} Number of taxa used for each pairwise contrast in the VCV analyses, given the number of taxa without inapplicable states.](Supplementary_materials/Online_appendices/PNGs/VCV_ntaxa.png)

![\label{VCV_regimes} Pairwise estimated rate covariance matrices across the five selective regimes, using only taxa with diet data. Covariances scaled to correlations. Selective regimes were mapped onto the tree (22 species with diet data) using a stochastic mapping of the feeding guilds. Tree is pruned to taxa with no inapplicable states for a given character pair. Not all regimes are represented in all contrasts. Question marks represent computationally singular contrasts.](Supplementary_materials/Online_appendices/PNGs/VCV_regimes.png){height=600px}

![\label{VCV_diffs} Scaled differences between the regime-specific covariance matrices in \@ref{VCV_regimes} and the whole tree covariance matrix.](Supplementary_materials/Online_appendices/PNGs/VCVdiffs.png){height=600px}

![\label{VCV_diffs_anc} Scaled differences between the regime-specific covariance matrices in \@ref{VCV_regimes} and the covariance matrices in their preceding regime, the large-crustacean specialist regime.](Supplementary_materials/Online_appendices/PNGs/VCVdiffs_anc.png)

### 2020-02-03 SysBio_Submission/Manuscript_postsubmission.Rmd
cleanup of previous submission materials and versions, organized to move forward with PlosBio/PNAS version

  Siphonophores have diverse predatory niches in the open ocean, ranging from mid-trophic small crustacean eaters to piscivorous super-carnivores. With the evolution of diversified prey type specializations comes the evolution of morphologies adapted to the challenges posed by different prey. The results presented here indicate that the associations found between siphonophore tentilla and their prey are a product of correlated evolution in highly integrated traits. While much of the literature focuses on how predatory generalists evolve into predatory specialists, in siphonophores we find predatory specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. Our extended morphological characterization shows that the relationships between form and ecology hold across a large set of taxa and characters, and can be used to generate hypotheses on the feeding habits of uncharacterized species. We conclude that the siphonophores were able to become abundant oceanic predators by occupying a variety of trophic niches facilitated by the evolution and diversification of extraordinary prey capture tools on their tentacles.

>>>>

  Siphonophores have diverse predatory niches in the open ocean, ranging from mid-trophic small crustacean eaters to piscivorous super-carnivores. With the evolution of diversified prey type specializations comes the evolution of morphologies adapted to the challenges posed by different prey. The results presented here indicate that the associations found between siphonophore tentilla and their prey are a product of correlated evolution in highly integrated traits. While much of the literature focuses on how predatory generalists evolve into predatory specialists, in siphonophores we find predatory specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. Our extended morphological characterization shows that the relationships between form and ecology hold across a large set of taxa and characters, and can be used to generate hypotheses on the feeding habits of uncharacterized species. We find that the evolutionary diversification of tools for prey capture contributed to the diversification of trophic interactions. Therefore, we identify organismal trait evolution as a key driver in the emergence of food web complexity.
